---
title: "R code not filtered"
author: "Rosalane"
date: '2022-06-29'
output: html_document
---

#Libraries needed
```{r}
library("phyloseq")
library("ggplot2")
library("ape")
library("gridExtra")
library("ggpubr")
library("ggsignif")
library(forcats)
library(vegan)
library(cowplot)
library(tidyverse)
```


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "C:/Users/kithanro/Documents/16s_Gala_postharvest_2021")
```

#Phyloseq object
```{r echo=TRUE}
#read in otu table
otu_table <- read.table("OTU_table.txt",
                        sep="\t",row.names = 1,header = TRUE) %>%
  as.matrix(otu_table,rownames=TRUE)

#read in taxonomy
taxonomy <- read_delim("taxonomy.csv", delim = ",", col_names = TRUE) %>% 
  column_to_rownames(var = "OTU_ID") %>%
  as.matrix()

#read in metadata
metadata=read.table("Sample_metadata.tsv", sep = "\t", header = TRUE,
                    row.names=1, na.strings = "") 
metadata
phy_tree = read_tree("tree.nwk")

#import as phyloseq objects
OTU = otu_table(otu_table,taxa_are_rows=TRUE)
TAX = tax_table(taxonomy)
META = sample_data(metadata)
#Physeq object
physeq <- phyloseq(OTU,TAX,META,phy_tree)
```


#Fitering Chloroplast and Mitocondria OTUs
```{r}
`%notin%` <- Negate(`%in%`)
# #Filtering samples *maybe only works on data frames
# physeq_filtered <- subset_taxa(physeq, Order %notin% "Chloroplast", 
#                                Family %notin% "Mitochondria")

#physeq <- physeq_filtered
physeq_filtered <- physeq %>% subset_taxa(Order != "Chloroplast") %>%
  subset_taxa(Family != "Mitohondria")

#check reads per sample
test_df <- as.data.frame(sample_sums(physeq_filtered)) %>% #sum reads per sample
  rownames_to_column(var = "Sample_ID") #format for plot

# make the plot
test_plot <- ggplot(test_df, aes(x = reorder(Sample_ID, -`sample_sums(physeq_filtered)`), #orders axis
                                 y = `sample_sums(physeq_filtered)`
                                 )) + 
  geom_col() + ylab("Reads per sample") + xlab("Sample_ID") + #clean up labels
  theme(axis.text.x = element_text(angle = -90, hjust=0))
test_plot

# more filtering to remove samples that didn't sequence well or OTUs that may not be real 
  #physeq_filtered2 <- physeq_filtered %>%
  #prune_samples(sample_sums(.) >= 1000, .) #remove samples with <=1000 reads per samples
  
# prune_taxa(taxa_sums(.) >= 2, .) # remove taxa with <=2 reads per OTU
  
# subset samples to each read numbers 
```


#Alpha diversity for Control
```{r fig.width=10, fig.height=40}
alpha_diversity <- estimate_richness(physeq_filtered, measures = c("Shannon", "Chao1"))
df_alpha <- data.frame(alpha_diversity, sample_data(physeq_filtered))
df <- reshape2::melt(df_alpha,
                     measure.var=c("Shannon","Chao1"),
                     id.vars=c("Timepoints", "Treatments", "Tissue_types"))
df2 <- df %>%
  #mutate(value = as.numeric(value)) %>% 
  #as.numeric(df$value)
  filter(Timepoints != "None") %>% 
  filter(Treatments != "ASM-treated") %>%
  filter(Treatments == "Control")
#selecting the rows corresponding to each index from the main dataframe
shannon = dplyr::filter(df2, variable == "Shannon")
chao = dplyr::filter(df2, variable == "Chao1")
```

##Alpha diversity plot
```{r}
library(ggsci)
#facet example
facet_plot <- df2 %>% 
  #filter(variable == "Shannon") %>%
  ggplot(aes(x = Tissue_types, y = value, fill = Timepoints)) +
  geom_boxplot(outlier.shape = NA) +
  geom_point(color="black", size=1, alpha=0.5, 
             position = position_jitterdodge(jitter.width = 0.1)) +
  facet_grid(variable~., scales = "free") +
  labs(y="Alpha Diversity Index", x="Tissue_types") +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
  theme_classic() + 
  theme(axis.title.x=element_blank()) + 
  ggtitle("Control-Not filtered") +
   scale_fill_aaas()
facet_plot
##################### Running box plot #############################
# shannon_plot <- shannon %>% #shannon_box, chao1_box
#   ggplot(aes(x = Treatments, y = value, fill = Timepoints)) +
#   geom_boxplot(outlier.shape = NA) +
#   geom_jitter(color="black", size=1, alpha=0.5) +
#   labs(y="Shannon", x="Treatments") +
#   annotate("text", x = 2, y = 3.2, label = "Kruskal-Wallis, p = 0.0097", colour = "black") +
#   theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
#   theme_classic() + 
#   theme(axis.title.x=element_blank()) + 
#    scale_fill_aaas()
# 
# shannon_plot
# chao_plot <- chao %>% #shannon_box, chao1_box
#   ggplot(aes(x = Treatments, y = value, fill = Timepoints)) +
#   geom_boxplot(outlier.shape = NA) +
#   geom_jitter(color="black", size=1, alpha=0.5) +
#   labs(y="Chao1", x="Treatments") +
#   annotate("text", x = 2, y = 130, label = "Kruskal-Wallis, p = 0.0032", colour = "black") +
#   theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
#   theme_classic() + 
#   theme(axis.title.x=element_blank()) + 
#    scale_fill_aaas()
#  chao_plot
```


#Alpha diversity for ASM-treated
```{r fig.width=10, fig.height=40}
alpha_diversity <- estimate_richness(physeq_filtered, measures = c("Shannon", "Chao1"))
df_alpha <- data.frame(alpha_diversity, sample_data(physeq_filtered))
df <- reshape2::melt(df_alpha,
                     measure.var=c("Shannon","Chao1"),
                     id.vars=c("Timepoints", "Treatments", "Tissue_types"))
df3 <- df %>%
  #mutate(value = as.numeric(value)) %>% 
  #as.numeric(df$value)
  filter(Timepoints != "None") %>% 
  filter(Treatments == "ASM-treated") %>%
  filter(Treatments != "Control")
#selecting the rows corresponding to each index from the main dataframe
shannon = dplyr::filter(df3, variable == "Shannon")
chao = dplyr::filter(df3, variable == "Chao1")
```

##Alpha diversity plot
```{r}
library(ggsci)
#facet example
facet_plot2 <- df3 %>% 
  #filter(variable == "Shannon") %>%
  ggplot(aes(x = Tissue_types, y = value, fill = Timepoints)) +
  geom_boxplot(outlier.shape = NA) +
  geom_point(color="black", size=1, alpha=0.5, 
             position = position_jitterdodge(jitter.width = 0.1)) +
  facet_grid(variable~., scales = "free") +
  labs(y="Alpha Diversity Index", x="Tissue_types") +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
  theme_classic() + 
  theme(axis.title.x=element_blank()) + 
  ggtitle("ASM-treated -Not filtered") +
   scale_fill_aaas()
facet_plot2
##################### Running box plot #############################
# shannon_plot <- shannon %>% #shannon_box, chao1_box
#   ggplot(aes(x = Treatments, y = value, fill = Timepoints)) +
#   geom_boxplot(outlier.shape = NA) +
#   geom_jitter(color="black", size=1, alpha=0.5) +
#   labs(y="Shannon", x="Treatments") +
#   annotate("text", x = 2, y = 3.2, label = "Kruskal-Wallis, p = 0.0097", colour = "black") +
#   theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
#   theme_classic() + 
#   theme(axis.title.x=element_blank()) + 
#    scale_fill_aaas()
# 
# shannon_plot
# chao_plot <- chao %>% #shannon_box, chao1_box
#   ggplot(aes(x = Treatments, y = value, fill = Timepoints)) +
#   geom_boxplot(outlier.shape = NA) +
#   geom_jitter(color="black", size=1, alpha=0.5) +
#   labs(y="Chao1", x="Treatments") +
#   annotate("text", x = 2, y = 130, label = "Kruskal-Wallis, p = 0.0032", colour = "black") +
#   theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
#   theme_classic() + 
#   theme(axis.title.x=element_blank()) + 
#    scale_fill_aaas()
#  chao_plot
```

# Phylum bar plots

### Control : not filtered

Format phyloseq object, remove chloroplasts, calculate mean abundace per group

```{r}
phylum <- physeq_filtered %>%
  subset_taxa(Family != "Mitochondria")%>%
  subset_taxa(Family != "Chloroplast")%>%
  subset_samples(Timepoints != "ASM-treated") %>%
  subset_samples(Timepoints != "None") %>%
  subset_samples(Treatments == "Control") %>%
  tax_glom("Phylum") %>%
  transform_sample_counts(function(x) { (x / sum(x) ) * 100} ) %>%
  psmelt() %>%

#make it after making the bar plot. Other because it is very less.
  #mutate(Phylum = recode(Phylum,
                        # "Armatimonadetes" = "Other",
                         #"candidate_division_WPS-1" = "Other",
                         #"Chlamydiae" = "Other", 
                         #"Deinococcus-Thermus" = "Other",
                         #"Fusobacteria" = "Other", 
                         #"Nitrospirae" = "Other", 
                         #"Parcubacteria" = "Other",
                         #"Synergistetes" = "Other" )) %>% 

  group_by(OTU, Tissue_types, Timepoints, Treatments, Phylum) %>%
  summarise(Mean_abund = round(mean(Abundance, na.rm = TRUE), digits = 2)) %>%
  ungroup() %>%
  as.data.frame()
```


Make the plot 

```{r}
plot_phylum <- phylum %>%
  ggplot(aes(x = Timepoints, y = Mean_abund, fill = Phylum, 
             Tissue_types)) + 
  geom_bar(stat = "identity") + 
  facet_grid(~Tissue_types) +
  # Clean up labels
  ylab("Relative Abundance") +
  ggtitle("Control- not filtered for low reads") +
  theme(axis.text.y = element_text(size = 10), #reduce y-axis size
        axis.text.x = element_text(angle = 90, #rotate tick labels
                                    vjust = 0.5, hjust = 1, #straigten up labels
                                    size = 10), #reduce x-axis size
        legend.text = element_text(size = 10)) #reduce legend size
  # customize plot colors
  # scale_fill_manual(name = "Phylum", #legend name
  #                   #breaks = phy_names13, #call legend labels
  #                   labels = phy_labels13, #cleaned up labels
  #                   values = phy_colors13, #custom colors 
  #                   guide = guide_legend(ncol = 1)) #force one column 
plot_phylum
```

# Phylum bar plots

### ASM-treated : No chloroplast and low reads filtered

Format phyloseq object, remove chloroplasts, calculate mean abundace per group

```{r}
phylum <- physeq_filtered %>%
  subset_taxa(Family != "Mitochondria")%>%
  subset_taxa(Family != "Chloroplast")%>%
  subset_samples(Timepoints != "Control") %>%
  subset_samples(Timepoints != "None") %>%
  subset_samples(Treatments == "ASM-treated") %>%
  tax_glom("Phylum") %>%
  transform_sample_counts(function(x) { (x / sum(x) ) * 100} ) %>%
  psmelt() %>%

#make it after making the bar plot. Other because it is very less.
  #mutate(Phylum = recode(Phylum,
                        # "Armatimonadetes" = "Other",
                         #"candidate_division_WPS-1" = "Other",
                         #"Chlamydiae" = "Other", 
                         #"Deinococcus-Thermus" = "Other",
                         #"Fusobacteria" = "Other", 
                         #"Nitrospirae" = "Other", 
                         #"Parcubacteria" = "Other",
                         #"Synergistetes" = "Other" )) %>% 

  group_by(OTU, Tissue_types, Timepoints, Treatments, Phylum) %>%
  summarise(Mean_abund = round(mean(Abundance, na.rm = TRUE), digits = 2)) %>%
  ungroup() %>%
  as.data.frame()
```


Make the plot 

```{r}
plot_phylum <- phylum %>%
  ggplot(aes(x = Timepoints, y = Mean_abund, fill = Phylum, 
             Tissue_types)) + 
  geom_bar(stat = "identity") + 
  facet_grid(~Tissue_types) +
  # Clean up labels
  ylab("Relative Abundance") +
  ggtitle("ASM-treated: not filtered for low reads") +
  theme(axis.text.y = element_text(size = 10), #reduce y-axis size
        axis.text.x = element_text(angle = 90, #rotate tick labels
                                    vjust = 0.5, hjust = 1, #straigten up labels
                                    size = 10), #reduce x-axis size
        legend.text = element_text(size = 10)) #reduce legend size
  # customize plot colors
  # scale_fill_manual(name = "Phylum", #legend name
  #                   #breaks = phy_names13, #call legend labels
  #                   labels = phy_labels13, #cleaned up labels
  #                   values = phy_colors13, #custom colors 
  #                   guide = guide_legend(ncol = 1)) #force one column 
plot_phylum
```

#Family for control- filtered

```{r}
Family <- physeq_filtered %>%
  subset_samples(Timepoints != "ASM-treated") %>%
  subset_samples(Timepoints != "None") %>%
  subset_samples(Treatments == "Control") %>%
  tax_glom("Family") %>%
  transform_sample_counts(function(x) { (x / sum(x) ) * 100} ) %>%
  psmelt() %>%

#make it after making the bar plot. Other because it is very less.
  #mutate(Phylum = recode(Phylum,
                        # "Armatimonadetes" = "Other",
                         #"candidate_division_WPS-1" = "Other",
                         #"Chlamydiae" = "Other", 
                         #"Deinococcus-Thermus" = "Other",
                         #"Fusobacteria" = "Other", 
                         #"Nitrospirae" = "Other", 
                         #"Parcubacteria" = "Other",
                         #"Synergistetes" = "Other" )) %>% 

  group_by(OTU, Tissue_types, Timepoints, Treatments, Family) %>%
  summarise(Mean_abund = round(mean(Abundance, na.rm = TRUE), digits = 2)) %>%
  ungroup() %>%
  as.data.frame()
```

Make the plot 

```{r}
plot_Family <- Family %>%
  ggplot(aes(x = Timepoints, y = Mean_abund, fill = Family, 
             Tissue_types)) + 
  geom_bar(stat = "identity") + 
  facet_grid(~Tissue_types) +
  # Clean up labels
  ylab("Relative Abundance") +
  ggtitle("Control- not filtered") +
  theme(axis.text.y = element_text(size = 10), #reduce y-axis size
        axis.text.x = element_text(angle = 90, #rotate tick labels
                                    vjust = 0.5, hjust = 1, #straigten up labels
                                    size = 10), #reduce x-axis size
        legend.text = element_text(size = 10)) #reduce legend size
  # customize plot colors
  # scale_fill_manual(name = "Phylum", #legend name
  #                   #breaks = phy_names13, #call legend labels
  #                   labels = phy_labels13, #cleaned up labels
  #                   values = phy_colors13, #custom colors 
  #                   guide = guide_legend(ncol = 1)) #force one column 
plot_Family
```

#Family for ASM-treated- not filtered

```{r}
Family <- physeq_filtered %>%
  subset_samples(Timepoints != "Control") %>%
  subset_samples(Timepoints != "None") %>%
  subset_samples(Treatments == "ASM-treated") %>%
  tax_glom("Family") %>%
  transform_sample_counts(function(x) { (x / sum(x) ) * 100} ) %>%
  psmelt() %>%

#make it after making the bar plot. Other because it is very less.
  #mutate(Phylum = recode(Phylum,
                        # "Armatimonadetes" = "Other",
                         #"candidate_division_WPS-1" = "Other",
                         #"Chlamydiae" = "Other", 
                         #"Deinococcus-Thermus" = "Other",
                         #"Fusobacteria" = "Other", 
                         #"Nitrospirae" = "Other", 
                         #"Parcubacteria" = "Other",
                         #"Synergistetes" = "Other" )) %>% 

  group_by(OTU, Tissue_types, Timepoints, Treatments, Family) %>%
  summarise(Mean_abund = round(mean(Abundance, na.rm = TRUE), digits = 2)) %>%
  ungroup() %>%
  as.data.frame()
```

Make the plot 

```{r}
plot_Family <- Family %>%
  ggplot(aes(x = Timepoints, y = Mean_abund, fill = Family, 
             Tissue_types)) + 
  geom_bar(stat = "identity") + 
  facet_grid(~Tissue_types) +
  # Clean up labels
  ylab("Relative Abundance") +
  ggtitle("ASM-treated - not filtered") +
  theme(axis.text.y = element_text(size = 10), #reduce y-axis size
        axis.text.x = element_text(angle = 90, #rotate tick labels
                                    vjust = 0.5, hjust = 1, #straigten up labels
                                    size = 10), #reduce x-axis size
        legend.text = element_text(size = 10)) #reduce legend size
  # customize plot colors
  # scale_fill_manual(name = "Phylum", #legend name
  #                   #breaks = phy_names13, #call legend labels
  #                   labels = phy_labels13, #cleaned up labels
  #                   values = phy_colors13, #custom colors 
  #                   guide = guide_legend(ncol = 1)) #force one column 
plot_Family
```