---
title: "Rot Microbiome 4"
author: "Kristi Gdanetz MacCready"
date: "4/6/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "~/Documents/Sundin_Lab/Rot_microbiome/")
```

# Load packages
```{r}
library(vegan)
library(tidyverse)
library(readxl)
library(agricolae)
library(phyloseq)
library(venn)
library(VennDiagram)
library(cowplot)
theme_set(theme_cowplot())
set.seed(7)
source("~/Documents/GitHub/DenefLab-MicrobeMiseq/R/miseqR.R")
```

# Import data 
Import object created in AlphaDiversity_analysis.Rmd 
```{r}
psq <- readRDS(file = "psq_object.RDS")
psq
# Summary statistics on read counts 
min(sample_sums(psq)) #23
mean(sample_sums(psq)) #15662
max(sample_sums(psq))  #28913
```

Import dataframe from Shared_Table_Formatting.Rmd
```{r}
tax <- readRDS(file = "tax.RDS")

shared_phylum<- readRDS(file = "shared_phylum.RDS")
```

# Other groups
Generate summaries of low abundance taxa to replace with "other" value in barplots
```{r eval=FALSE, include=FALSE}
# # occurences of each phylum in all data 
# counts1 <- shared_phylum %>%
#   column_to_rownames(var = "Sample_ID") %>%
#   colSums() %>%
#   as.data.frame() 

# Occur in less than 10 samples:
#Chlamydiae, Fusobacteria, Synergistetes, Nitrospirae, candidate_division_WPS-1, Armatimonadetes, Parcubacteria, Deinococcus-Thermus
# not in all, keep color codes for these: Chloroflexi, Latescibacteria

# phylum to recode as "other"; can remove colors for these groups
new_phyla <- tax %>%
  mutate(Phylum = recode(Phylum,
                         "Armatimonadetes" = "Other",
                         "candidate_division_WPS-1" = "Other",
                         "Chlamydiae" = "Other", 
                         "Deinococcus-Thermus" = "Other",
                         "Fusobacteria" = "Other", 
                         "Nitrospirae" = "Other", 
                         "Parcubacteria" = "Other",
                         "Synergistetes" = "Other" ))
```

Total relative abundance in cultivar timecourse
```{r}
counts2 <- phylum %>%
  group_by(Phylum) %>%
  summarise(Abund_sum = sum(Mean_abund))
# less than 0.10, replace with "Other": 
#Chlamydiae, Fusobacteria, Nitrospirae, Synergistetes, Armatimonadetes, candidate_division_WPS-1, Parcubacteria, Deinococcus-Thermus, 
# not in all, keep color codes for these: Chloroflexi, Latescibacteria, 

new_phyla2 <- tax %>%
  mutate(Phylum = recode(Phylum,
                         "Armatimonadetes" = "Other",
                         "candidate_division_WPS-1" = "Other",
                         "Chlamydiae" = "Other", 
                         "Deinococcus-Thermus" = "Other",
                         "Fusobacteria" = "Other", 
                         "Nitrospirae" = "Other", 
                         "Parcubacteria" = "Other",
                         "Synergistetes" = "Other" ,
                         "Chloroflexi" = "Other",
                         "Latescibacteria" = "Other"))
```
total relative abundace in chitosan dataset
```{r}
counts3 <- phylum_chito %>%
  group_by(Phylum) %>%
  summarise(Abund_sum = sum(Mean_abund)) 
# less than 0.10, replace with "Other": 
#Armatimonadetes, candidate_division_WPS-1, Chlamydiae, Deinococcus-Thermus, Fusobacteria, Nitrospirae, Parcubacteria, Synergistetes
# not in all, keep color codes for these: Abditibacteriota, Chloroflexi, Gemmatimonadetes, 

new_phyla3 <- tax %>%
  mutate(Phylum = recode(Phylum,
                         "Armatimonadetes" = "Other",
                         "candidate_division_WPS-1" = "Other",
                         "Chlamydiae" = "Other", 
                         "Deinococcus-Thermus" = "Other",
                         "Fusobacteria" = "Other", 
                         "Nitrospirae" = "Other", 
                         "Parcubacteria" = "Other",
                         "Synergistetes" = "Other",
                         "Abditibacteriota" = "Other",
                         "Chloroflexi" = "Other", 
                         "Gemmatimonadetes" = "Other"))
```
total relative abundace in rotten samples
```{r}
counts4 <- phylum_rot %>%
  group_by(Phylum) %>%
  summarise(Abund_sum = sum(Mean_abund)) 
# less than 0.10, replace with "Other":
#Chlamydiae,Fusobacteria,Nitrospirae,Synergistetes,Armatimonadetes,candidate_division_WPS-1, Parcubacteria, Deinococcus-Thermus
# not in all, keep color codes for these: Latescibacteria, 

new_phyla4 <- tax %>%
  mutate(Phylum = recode(Phylum,
                         "Armatimonadetes" = "Other",
                         "candidate_division_WPS-1" = "Other",
                         "Chlamydiae" = "Other", 
                         "Deinococcus-Thermus" = "Other",
                         "Fusobacteria" = "Other", 
                         "Nitrospirae" = "Other", 
                         "Parcubacteria" = "Other",
                         "Synergistetes" = "Other",
                         "Latescibacteria" = "Other"))
```

# Color palette
### Phyla 21 colors 
generated from https://medialab.github.io/iwanthue/
options: pimp, 21, force vector, sort by diff
```{r eval=FALSE, include=FALSE}
phy_colors21 <- c("#6c679a", "#93b824", "#af00c9", "#017100", "#8300a9",
                "#2fc185", "#3f1ba2", "#ab8800", "#302a95", "#ff7823",
                "#006dc5", "#b80018", "#44b0ff", "#754700", "#910075",
                "#9bb455", "#7f002b", "#00471e", "#ff637f", "#73733e",
                "#d5a078", "gray", "black")

phy_labels21 <- c("Abditibacteriota", "Acidobacteria", "Actinobacteria", "Armatimonadetes", "Unclassified Bacteria",
                "Bacteroidetes", "Candidate Division WPS-1", "Chlamydiae", "Chloroflexi", "Chloroplast/Cyanobacteria",
                "Deinococcus-Thermus", "Firmicutes", "Fusobacteria", "Gemmatimonadetes", "Latescibacteria",
                "Nitrospirae", "Parcubacteria", "Planctomycetes", "Proteobacteria", "Synergistetes",
                "Verrucomicrobia")

phy_names21 <- c("Abditibacteriota", "Acidobacteria", "Actinobacteria", "Armatimonadetes", "Bacteria_unclassified",
                "Bacteroidetes", "candidate_division_WPS-1", "Chlamydiae", "Chloroflexi", "Cyanobacteria/Chloroplast",
                "Deinococcus-Thermus", "Firmicutes", "Fusobacteria", "Gemmatimonadetes", "Latescibacteria",
                "Nitrospirae", "Parcubacteria", "Planctomycetes", "Proteobacteria", "Synergistetes",
                "Verrucomicrobia")
```

### Phyla 13 colors
generated from https://medialab.github.io/iwanthue/ 
options: pimp, 13, force vector, sort by chroma
Added neutral gray to color list; will use for "other" group
```{r}
## original color order:
# phy_colors13 <- c("#007aa8","#662052","#e190d3","#019863","#ff7a86",
#                   "#926800","#94b747","#891700","#498200","#0389f8",
#                   "#cb0050","#a775ff","#ff15b1", "#8e918f")

# actinobacteria = bubble gum pink; bacteroidetes = salmon pink; unclassified bacteria = dark green; firmicutes = maroon; proteobacteria = lt purple;  
# switch proteobacteria "#a775ff" and abditibacteriota "#007aa8"
# actinobacteria "#662052" and gemmatimonadetes "#498200"
# firmicutes "#891700" and plactomycetes "#cb0050" 
# actinobacteria "#e190d3" and latescibacteria "#0389f8"

# updated color order: 
phy_colors13 <- c("#a775ff","#498200","#0389f8","#019863","#ff7a86",
                  "#926800","#94b747","#cb0050","#662052","#e190d3",
                  "#891700","#007aa8","#ff15b1", "#8e918f")

phy_labels13 <- c("Abditibacteriota", "Acidobacteria", "Actinobacteria", "Unclassified Bacteria", "Bacteroidetes",
                 "Chloroflexi", "Chloroplast/Cyanobacteria", "Firmicutes", "Gemmatimonadetes", "Latescibacteria",
                 "Planctomycetes", "Proteobacteria", "Verrucomicrobia", "Other")

phy_names13 <- c("Abditibacteriota", "Acidobacteria", "Actinobacteria", "Bacteria_unclassified", "Bacteroidetes",
                 "Chloroflexi", "Cyanobacteria/Chloroplast", "Firmicutes", "Gemmatimonadetes", "Latescibacteria",
                 "Planctomycetes", "Proteobacteria", "Verrucomicrobia", "Other")
```

# Phylum bar plots

### No chloroplasts
Format phyloseq object, remove chloroplasts, calculate mean abundace per group
```{r}
phylum <- psq %>%
  subset_taxa(Phylum != "Cyanobacteria/Chloroplast") %>%
  subset_samples(Timepoint != "Rotten") %>%
  subset_samples(Treatment == "Untreated") %>%
  tax_glom("Phylum") %>%
  transform_sample_counts(function(x) { (x / sum(x) ) * 100} ) %>%
  psmelt() %>%

#make it after making the bar plot. Other because it is very less.
  #mutate(Phylum = recode(Phylum,
                        # "Armatimonadetes" = "Other",
                         #"candidate_division_WPS-1" = "Other",
                         #"Chlamydiae" = "Other", 
                         #"Deinococcus-Thermus" = "Other",
                         #"Fusobacteria" = "Other", 
                         #"Nitrospirae" = "Other", 
                         #"Parcubacteria" = "Other",
                         #"Synergistetes" = "Other" )) %>% 

  group_by(OTU, Tissue, Timepoint, Cultivar, Treatment, Phylum) %>%
  summarise(Mean_abund = round(mean(Abundance, na.rm = TRUE), digits = 2)) %>%
  ungroup() %>%
  as.data.frame()
```
Make the plot 
```{r}
plot_phylum <- phylum %>%
  ggplot(aes(x = Timepoint, y = Mean_abund, fill = Phylum, 
             Tissue, Tree, Cultivar)) + 
  geom_bar(stat = "identity") + 
  facet_grid(Cultivar~Tissue) +
  # Clean up labels
  ylab("Relative Abundance") +
  theme(axis.text.y = element_text(size = 10), #reduce y-axis size
        axis.text.x = element_text(angle = 90, #rotate tick labels
                                    vjust = 0.5, hjust = 1, #straigten up labels
                                    size = 10), #reduce x-axis size
        legend.text = element_text(size = 10)) + #reduce legend size
  # customize plot colors
  scale_fill_manual(name = "Phylum", #legend name
                    breaks = phy_names13, #call legend labels
                    labels = phy_labels13, #cleaned up labels
                    values = phy_colors13, #custom colors 
                    guide = guide_legend(ncol = 1)) #force one column 
plot_phylum
```

### With chloropasts
Repeat, but keep chloroplasts
```{r}
phylum_chloro <- psq %>%
  # subset for samples of interest
  subset_samples(Timepoint != "Rotten") %>%
  subset_samples(Treatment == "Untreated") %>%
  # combine OTUs based on phylum assignments
  tax_glom("Phylum") %>%
  # convert to relative abundance
  transform_sample_counts(function(x) { (x / sum(x) ) * 100} ) %>%
  # make data frame
  psmelt() %>%
  # replace low abundance taxa with "other"
  mutate(Phylum = recode(Phylum,
                         "Armatimonadetes" = "Other",
                         "candidate_division_WPS-1" = "Other",
                         "Chlamydiae" = "Other", 
                         "Deinococcus-Thermus" = "Other",
                         "Fusobacteria" = "Other", 
                         "Nitrospirae" = "Other", 
                         "Parcubacteria" = "Other",
                         "Synergistetes" = "Other" )) %>% 
  # calculate mean relative abunance of replicates 
  group_by(OTU, Tissue, Timepoint, Cultivar, Treatment, Phylum) %>%
  summarise(Mean_abund = round(mean(Abundance, na.rm = TRUE), digits = 2)) %>%
  ungroup() %>%
  as.data.frame()
```

```{r}
plot_chloro <- phylum_chloro %>%
  ggplot(aes(x = Timepoint, y = Mean_abund, fill = Phylum, 
             Tissue, Tree, Cultivar)) + 
  geom_bar(stat = "identity") + 
  facet_grid(Cultivar~Tissue) +
  # format legend
  guides(fill = guide_legend(ncol = 1)) +  
  # Clean up labels
  ylab("Relative Abundance") +
  theme(axis.text.y = element_text(size = 10), #reduce y-axis size
        axis.text.x = element_text(angle = 90, #rotate tick labels
                                    vjust = 0.5, hjust = 1, #straigten up labels
                                    size = 10), #reduce x-axis size
        legend.text = element_text(size = 10)) + #reduce legend size
  # customize plot colors
  scale_fill_manual(name = "Phylum", #legend name
                    breaks = phy_names13, #call legend labels
                    labels = phy_labels13, #cleaned up labels
                    values = phy_colors13, #custom colors 
                    guide = guide_legend(ncol = 1)) #force one column 
plot_chloro
```

### Rotten
Format phyloseq object, remove chloroplasts, include rotten sames as timepoint, calculate mean abundace per group
```{r}
phylum_rot <- psq %>%
  subset_taxa(Phylum != "Cyanobacteria/Chloroplast") %>%
  subset_samples(Treatment == "Untreated") %>%
  tax_glom("Phylum") %>%
  transform_sample_counts(function(x) { (x / sum(x) ) * 100} ) %>%
  psmelt() %>%
  mutate(Phylum = recode(Phylum,
                         "Armatimonadetes" = "Other",
                         "candidate_division_WPS-1" = "Other",
                         "Chlamydiae" = "Other", 
                         "Deinococcus-Thermus" = "Other",
                         "Fusobacteria" = "Other", 
                         "Nitrospirae" = "Other", 
                         "Parcubacteria" = "Other",
                         "Synergistetes" = "Other",
                         "Latescibacteria" = "Other")) %>% 
  group_by(OTU, Tissue, Timepoint, Cultivar, Treatment, Phylum) %>%
  summarise(Mean_abund = round(mean(Abundance, na.rm = TRUE), digits = 2)) %>%
  ungroup() %>% 
  as.data.frame()
```
Make the plot 
```{r}
plot_rot <- phylum_rot %>%
  ggplot(aes(x = Timepoint, y = Mean_abund, fill = Phylum, 
             Tissue, Cultivar)) + 
  geom_bar(stat = "identity") + 
  facet_grid(Cultivar~Tissue) +
  # format legend
  guides(fill = guide_legend(ncol = 1)) +  
  # Clean up labels
  ylab("Relative Abundance") +
  theme(axis.text.y = element_text(size = 10), #reduce y-axis size
        axis.text.x = element_text(angle = 90, #rotate tick labels
                                    vjust = 0.5, hjust = 1, #straigten up labels
                                    size = 10), #reduce x-axis size
        legend.text = element_text(size = 10)) + #reduce legend size 
  # customize plot colors
  scale_fill_manual(name = "Phylum", #legend name
                    breaks = phy_names13, #call legend labels
                    labels = phy_labels13, #cleaned up labels
                    values = phy_colors13, #custom colors 
                    guide = guide_legend(ncol = 1)) #force one column 
plot_rot
```

### Chitosan
Subset for chitosan and mock treated samples, format phyloseq object
```{r}
phylum_chito <- psq %>%
  subset_taxa(Phylum != "Cyanobacteria/Chloroplast") %>%
  subset_samples(Cultivar == "Jonathan") %>%
  subset_samples(Treatment == "Chitosan" | Treatment == "Mock") %>%
  tax_glom("Phylum") %>%
  transform_sample_counts(function(x) { (x / sum(x) ) * 100} ) %>%
  psmelt() %>%
  mutate(Phylum = recode(Phylum,
                         "Armatimonadetes" = "Other",
                         "candidate_division_WPS-1" = "Other",
                         "Chlamydiae" = "Other", 
                         "Deinococcus-Thermus" = "Other",
                         "Fusobacteria" = "Other", 
                         "Nitrospirae" = "Other", 
                         "Parcubacteria" = "Other",
                         "Synergistetes" = "Other",
                         "Abditibacteriota" = "Other",
                         "Chloroflexi" = "Other", 
                         "Gemmatimonadetes" = "Other")) %>% 
  group_by(OTU, Tissue, Timepoint, Cultivar, Treatment, Phylum) %>%
  summarise(Mean_abund = round(mean(Abundance, na.rm = TRUE), digits = 2)) %>%
  ungroup() %>% 
  as.data.frame()
```
Make the plot
```{r}
plot_chito <- phylum_chito %>%
  ggplot(aes(x = Timepoint, y = Mean_abund, fill = Phylum, 
             Tissue, Treatment)) + 
  geom_bar(stat = "identity") + 
  facet_grid(Treatment~Tissue) +
  # format legend
  guides(fill = guide_legend(ncol = 1)) +  
  # Clean up labels
  ylab("Relative Abundance") +
  theme(axis.text.y = element_text(size = 10), #reduce y-axis size
        axis.text.x = element_text(angle = 90, #rotate tick labels
                                    vjust = 0.5, hjust = 1, #straigten up labels
                                    size = 10), #reduce x-axis size
        legend.text = element_text(size = 10)) + #reduce legend size 
  # customize plot colors
  scale_fill_manual(name = "Phylum", #legend name
                    breaks = phy_names13, #call legend labels
                    labels = phy_labels13, #cleaned up labels
                    values = phy_colors13, #custom colors 
                    guide = guide_legend(ncol = 1)) #force one column 
plot_chito
```

# Proteobacteria by genus
Format phyloseq object, filter for proteobacteria, remove chloroplasts, calculate mean abundace per group
```{r}
proteobacteria <- psq %>%
  subset_taxa(Phylum != "Cyanobacteria/Chloroplast") %>%
  subset_taxa(Phylum == "Proteobacteria") %>%
  subset_samples(Treatment == "Untreated") %>%
  tax_glom("Genus") %>%
  transform_sample_counts(function(x) { (x / sum(x) ) * 100} ) %>%
  psmelt() %>%
  group_by(OTU, Tissue, Timepoint, Cultivar, Treatment, Genus) %>%
  summarise(Mean_abund = round(mean(Abundance, na.rm = TRUE), digits = 2)) %>%
  ungroup() %>%
  as.data.frame() %>%
  #filter(Mean_abund >= 1.00) %>%
  mutate(Genus = recode(Genus, 
                        Acetobacteraceae_unclassified = "Other",
                        Acidisoma = "Other",
                        Acidocella = "Other",
                        Acinetobacter = "Other",
                        Alcaligenaceae_unclassified = "Other",
                        Altererythrobacter = "Other",
                        Aminobacter = "Other",
                        Aquabacterium = "Other",
                        Archangium = "Other",
                        Arsenophonus = "Other",
                        Aurantimonadaceae_unclassified = "Other",
                        Aureimonas = "Other",
                        Azohydromonas = "Other",
                        Bacteriovoracales_unclassified = "Other",
                        Beijerinckiaceae_unclassified = "Other",
                        Belnapia = "Other",
                        Bosea = "Other",
                        Bradyrhizobiaceae_unclassified = "Other",
                        Bradyrhizobium = "Other",
                        Brevundimonas = "Other",
                        Brucella = "Other",
                        Buchnera = "Other",
                        Burkholderiaceae_unclassified = "Other",
                        Burkholderiales_unclassified = "Other",
                        Caulobacter = "Other",
                        Caulobacteraceae_unclassified = "Other",
                        Cellvibrio = "Other",
                        Comamonas = "Other",
                        Croceibacterium = "Other",
                        Cupriavidus = "Other",
                        Curvibacter = "Other",
                        Cystobacter = "Other",
                        Cystobacteraceae_unclassified = "Other",
                        Dechloromonas = "Other",
                        Delftia = "Other",
                        Deltaproteobacteria_unclassified = "Other",
                        Devosia = "Other",
                        Dokdonella = "Other",
                        Duganella = "Other",
                        Duodenibacillus = "Other",
                        Dyella = "Other",
                        Enhydrobacter = "Other",
                        Erythrobacteraceae_unclassified = "Other",
                        Frateuria = "Other",
                        Gammaproteobacteria_unclassified = "Other",
                        Gluconobacter = "Other",
                        Haematobacter = "Other",
                        Halomonas = "Other",
                        Hyphomicrobiaceae_unclassified = "Other",
                        Kingella = "Other",
                        Klebsiella = "Other",
                        Komagataeibacter = "Other",
                        Kosakonia = "Other",
                        Labilithrix = "Other",
                        Legionella = "Other",
                        Luteibacter = "Other",
                        Mameliella = "Other",
                        Mesorhizobium = "Other",
                        Methylobacteriaceae_unclassified = "Other",
                        Methylorubrum = "Other",
                        Microvirga = "Other",
                        Moraxellaceae_unclassified = "Other",
                        Myxococcales_unclassified = "Other",
                        Neorhizobium = "Other",
                        Nevskia = "Other",
                        Noviherbaspirillum = "Other",
                        Novosphingobium = "Other",
                        Oligoflexales_unclassified = "Other",
                        Oxalobacteraceae_unclassified = "Other",
                        Pantoea = "Other",
                        Paracoccus = "Other",
                        Pararhizobium = "Other",
                        Pasteurellaceae_unclassified = "Other",
                        Pectobacteriaceae_unclassified = "Other",
                        Peredibacter = "Other",
                        Phenylobacterium = "Other",
                        Phreatobacter = "Other",
                        Phyllobacterium = "Other",
                        Piscinibacter = "Other",
                        Polaromonas = "Other",
                        Polyangiaceae_unclassified = "Other",
                        Polyangium = "Other",
                        Proteus = "Other",
                        Pseudolabrys = "Other",
                        Psychroglaciecola = "Other",
                        Ramlibacter = "Other",
                        Reyranella = "Other",
                        Rhizobiaceae_unclassified = "Other",
                        Rhizobium = "Other",
                        Rhodanobacteraceae_unclassified = "Other",
                        Rhodobacteraceae_unclassified = "Other",
                        Rhodoplanes = "Other",
                        Rhodospirillales_unclassified = "Other",
                        Rickettsia = "Other",
                        Roseiarcus = "Other",
                        Roseivivax = "Other",
                        Roseomonas = "Other",
                        Rubellimicrobium = "Other",
                        Skermanella = "Other",
                        Sphingobium = "Other",
                        Sphingomonadales_unclassified = "Other",
                        Stenotrophomonas = "Other",
                        Tepidimonas = "Other",
                        Tepidiphilus = "Other",
                        Thauera = "Other",
                        Undibacterium = "Other",
                        Vampirovibrio = "Other",
                        Variovorax = "Other",
                        Wohlfahrtiimonas_unclassified = "Other",
                        Xanthomonas = "Other",
                        Yersiniaceae_unclassified = "Other"
                        )) %>%
  arrange(desc(Mean_abund))
```

Make the plot 
```{r}
plot_proteo <- proteobacteria %>%
  ggplot(aes(x = Timepoint, y = Mean_abund, fill = Genus, 
             Tissue, Tree, Cultivar)) + 
  geom_bar(stat = "identity") + 
  facet_grid(Cultivar~Tissue) +
  # Clean up labels
  ylab("Proteobacteria\nRelative Abundance") +
  theme(axis.text.y = element_text(size = 10), #reduce y-axis size
        axis.text.x = element_text(angle = 90, #rotate tick labels
                                    vjust = 0.5, hjust = 1, #straigten up labels
                                    size = 10), #reduce x-axis size
        legend.text = element_text(size = 10)) + #reduce legend size
  # customize plot colors
  scale_fill_manual(values = c("#4cbfcf", "#439774", "#6589cc", 
                               "#96b46e", "#9d4664",
                                "#906c2d", "#db8bc2", "#c8ab60", 
                               "#e17e79", "#707a2b",
                                "#a05032", "#8b60a9", "#3f7933", "#4fc284", "#dd894d",
                                "#d2468f", "#d23c55", "#d49c2b", 
                               "#a7b635", "#6f67d7",
                                "#5db746", "#bd56c3", "#d4502b", "gray"), 
                    breaks = c("Achromobacter", "Alphaproteobacteria_unclassified","Betaproteobacteria_unclassified",
                               "Burkholderia",  "Comamonadaceae_unclassified", 
                               "Enterobacterales_unclassified","Enterobacteriaceae_unclassified", "Escherichia/Shigella", 
                               "Haemophilus","Massilia",
                               "Methylobacterium","Neisseria", "Paraburkholderia","Pelomicrobium_unclassified","Pelomonas",
                               "Phyllobacteriaceae_unclassified","Pseudomonadaceae_unclassified","Proteobacteria_unclassified",
                               "Pseudomonas","Ralstonia",
                               "Rhizobiales_unclassified", "Sphingomonadaceae_unclassified",  "Sphingomonas",  "Other"),
                    labels = c("Achromobacter", "Alphaproteobacteria","Betaproteobacteria","Burkholderia","Comamonadaceae", 
                               "Enterobacterales","Enterobacteriaceae", "Escherichia/Shigella", "Haemophilus","Massilia",
                               "Methylobacterium","Neisseria", "Paraburkholderia","Pelomicrobium","Pelomonas",
                               "Phyllobacteriaceae","Pseudomonadaceae","Proteobacteria","Pseudomonas","Ralstonia",
                               "Rhizobiales", "Sphingomonadaceae",  "Sphingomonas",  "Other"),
                    guide = guide_legend(ncol = 2)) # 
plot_proteo
```

# Bacteroidetes by genus
Format data
```{r}
bacteroidetes <- psq %>%
  subset_taxa(Phylum != "Cyanobacteria/Chloroplast") %>%
  subset_taxa(Phylum == "Bacteroidetes") %>%
  subset_samples(Treatment == "Untreated") %>%
  tax_glom("Genus") %>%
  transform_sample_counts(function(x) { (x / sum(x) ) * 100} ) %>%
  psmelt() %>%
  group_by(OTU, Tissue, Timepoint, Cultivar, Treatment, Genus) %>%
  summarise(Mean_abund = round(mean(Abundance, na.rm = TRUE), digits = 2)) %>%
  ungroup() %>%
  as.data.frame() %>% 
  #filter(Mean_abund >= 1.00) %>% arrange(Genus, Cultivar, Tissue, Timepoint)
  mutate(Genus = recode(Genus, 
                      Adhaeribacter = "Other",
                      Adhaeribacter = "Other",
                      Adhaeribacter = "Other",
                      Bacteroides = "Other",
                      Belliella = "Other",
                      Cloacibacterium = "Other",
                      Cytophagales_unclassified = "Other",
                      Dysgonomonadaceae_unclassified = "Other",
                      Ferruginibacter = "Other",
                      Flavisolibacter = "Other",
                      Flavobacteriales_unclassified = "Other",
                      Flavobacterium = "Other",
                      Fulvivirgaceae_unclassified = "Other",
                      Hydrotalea = "Other",
                      Parabacteroides = "Other",
                      Parafilimonas = "Other",
                      Porphyromonadaceae_unclassified = "Other",
                      Porphyromonas = "Other",
                      Prevotella = "Other",
                      Prevotellaceae_unclassified = "Other",
                      Raineya = "Other",
                      Segetibacter = "Other")) %>%
  arrange(desc(Mean_abund))
```
Make the plot
```{r}
plot_bacter <- bacteroidetes %>%
  ggplot(aes(x = Timepoint, y = Mean_abund, fill = Genus, 
             Tissue, Tree, Cultivar)) + 
  geom_bar(stat = "identity") + 
  facet_grid(Cultivar~Tissue) +
  # Clean up labels
  ylab("Bacteroidetes\nRelative Abundance") +
  theme(axis.text.y = element_text(size = 10), #reduce y-axis size
        axis.text.x = element_text(angle = 90, #rotate tick labels
                                    vjust = 0.5, hjust = 1, #straighten up labels
                                    size = 10), #reduce x-axis size
        legend.text = element_text(size = 10)) + #reduce legend size
  # customize plot colors
  scale_fill_manual(values = c("#4cbfcf", "#439774", 
                               "#6589cc",  "#96b46e", 
                               "#9d4664", "#906c2d", 
                               "#db8bc2", "#c8ab60", 
                               "#e17e79", 
                               "gray",
                               "#707a2b", "#d2468f", 
                               "#8b60a9", "#3f7933",  "#4fc284"
                               #"#dd894d",  "#d23c55", "#d49c2b",  "#a05032",
                               #"#a7b635", "#6f67d7", "#5db746", "#bd56c3", "#d4502b", 
                               ), 
                    breaks = c("Bacteroidales_unclassified", "Bacteroidetes_unclassified",
                               "Chitinophagaceae_unclassified", "Chryseobacterium", 
                               "Dyadobacter", "Dysgonomonas", 
                               "Hydrobacter", "Hymenobacter", 
                               "Mucilaginibacter",
                               "Other", 
                               "Pedobacter", "Sediminibacterium",
                               "Sphingobacteriaceae_unclassified", "Spirosoma", "Terrimonas"),
                    labels = c("Bacteroidales unclassified", "Bacteroidetes unclassified",
                               "Chitinophagaceae unclassified","Chryseobacterium", 
                               "Dyadobacter", "Dysgonomonas", 
                               "Hydrobacter", "Hymenobacter", 
                               "Mucilaginibacter",
                               "Other", 
                               "Pedobacter", "Sediminibacterium",
                               "Sphingobacteriaceae unclassified", "Spirosoma", "Terrimonas"),
                    guide = guide_legend(ncol = 1)) # 
plot_bacter
```

# Acidobacteria by genus
Format data
```{r}
acidobact <- psq %>%
  subset_taxa(Phylum != "Cyanobacteria/Chloroplast") %>%
  subset_taxa(Phylum == "Acidobacteria") %>%
  subset_samples(Treatment == "Untreated") %>%
  tax_glom("Genus") %>%
  transform_sample_counts(function(x) { (x / sum(x) ) * 100} ) %>%
  psmelt() %>%
  group_by(OTU, Tissue, Timepoint, Cultivar, Treatment, Genus) %>%
  summarise(Mean_abund = round(mean(Abundance, na.rm = TRUE), digits = 2)) %>%
  ungroup() %>%
  as.data.frame() %>% 
  #filter(Mean_abund >= 1.00) %>% arrange(Genus, Cultivar, Tissue, Timepoint)
  # mutate(Genus = recode(Genus, 
  #                     Acidobacteria_Gp1_unclassified = "Other",
  #                     Acidobacteria_Gp3_unclassified = "Other",
  #                     Candidatus_Solibacter_unclassified = "Other",
  #                     Gp17_unclassified = "Other",
  #                     Gp5_unclassified = "Other",
  #                     Gp7_unclassified = "Other",
  #                     Terriglobus_unclassified = "Other")) %>%
  arrange(desc(Mean_abund))
```
Make the plot
```{r}
plot_acido <- acidobact %>%
  ggplot(aes(x = Timepoint, y = Mean_abund, fill = Genus, 
             Tissue, Tree, Cultivar)) + 
  geom_bar(stat = "identity") + 
  facet_grid(Cultivar~Tissue) +
  # Clean up labels
  ylab("Acidobacteria\nRelative Abundance") +
  theme(axis.text.y = element_text(size = 10), #reduce y-axis size
        axis.text.x = element_text(angle = 90, #rotate tick labels
                                    vjust = 0.5, hjust = 1, #straigten up labels
                                    size = 10), #reduce x-axis size
        legend.text = element_text(size = 10)) + #reduce legend size
  # customize plot colors
  scale_fill_manual(values = c("#4cbfcf", "#439774", "#6589cc", 
                               "#96b46e", "#9d4664",
                                "#906c2d", "#db8bc2", "#c8ab60", 
                               "#e17e79", "#707a2b",
                                "#a05032", "#8b60a9", "#3f7933", "#4fc284", "#dd894d",
                                "#d2468f", "#d23c55", "#d49c2b", 
                               "#a7b635", "#6f67d7",
                                "#5db746", "#bd56c3", "#d4502b", "gray"), 
                    #breaks = c(),
                    #labels = c(),
                    guide = guide_legend(ncol = 1)) # 
plot_acido
```

# Save bar pots
```{r}
# get legend with all phylum + color combinations 
bar_leg <- get_legend(plot_chloro)

# untreated, no rotten, no chloroplasts
save_plot(plot = plot_grid((plot_phylum + theme(legend.position = "none")), 
                           bar_leg, nrow = 1, ncol = 2, rel_widths = c(2, 0.75)), 
          base_height = 5, base_width = 8,
          filename = "Figures/Barplot_phylum_NoChloroplast.pdf") 

# untreated, no rotten, WITH chloroplasts
# cloroplast = lime green
save_plot(plot = plot_grid((plot_chloro + theme(legend.position = "none")), 
                           bar_leg, nrow = 1, ncol = 2, rel_widths = c(2, 0.75)), 
          base_height = 5, base_width = 8,
          filename = "Figures/Barplot_phylum_chloroplast.pdf") 

# untreated, including rotten, no chloroplasts
save_plot(plot = plot_grid((plot_rot + theme(legend.position = "none")), 
                           bar_leg, nrow = 1, ncol = 2, rel_widths = c(2, 0.75)), 
          base_height = 5, base_width = 8,
          filename = "Figures/Barplot_phylum_rotten.pdf") 

# chitosan treatments, jonathan only
save_plot(plot = plot_grid((plot_chito + theme(legend.position = "none")), 
                           bar_leg, nrow = 1, ncol = 2, rel_widths = c(2, 0.75)), 
          base_height = 5, base_width = 8,
          filename = "Figures/Barplot_phylum_chitosan.pdf")

# proteobateria by genera
save_plot(plot = plot_proteo, base_height = 5, base_width = 8,
          filename = "Figures/Barplot_proteobacteria_genus.pdf")
# bacteroidetes by genera
save_plot(plot = plot_bacter, base_height = 5, base_width = 8,
          filename = "Figures/Barplot_bacteroidetes_genus.pdf")
# acidobacteria by genera
save_plot(plot = plot_acido, base_height = 5, base_width = 8,
          filename = "Figures/Barplot_acidobacteria_genus.pdf")
```


# Export
```{r}

```

-----
end